---
title: "my_test_space"
author: "Cristina Guzman-Moumtzis"
date: "2023-07-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ~~~~
## Raw Data Processing
## ~~~~

## Set Up
This script is used as a test space to follow along the tutorials posted on MetaboAnalyst's website. 
The commands and data are all sourced from "2.1.2 Download IBD Example QC Data" on the tutorial tab

```{r folder set up, echo=FALSE}
## Setting the data depositing folder
data_folder_Sample <- "~/Data_IBD"
data_folder_QC <- "~/QC_IBD"  
```

## Downloading Sample Data
Access the sample data provided from this tutorial from Google Drive
Data is saved as a zipped folder locally

```{r data import}
# Use Google API for data downloading here. 
# Please "install.packages('googledrive')" and "install.packages('httpuv')"first.
library(googledrive);
temp <- tempfile(fileext = ".zip")
# Please authorize your google account to access the data
dl <- drive_download(
  as_id("10DBpPEWy2cZyXvmlLOIfpwqQYwFplKYK"), path = temp, overwrite = TRUE)
# Setting your own date file folder
out <- unzip(temp, exdir = data_folder_QC)
# Date files for parameters optimization are deposited below
out
```

## Unzip and save data
Previously loaded data is unzipped and saved locally

```{r data file}
temp <- tempfile(fileext = ".zip")
dl <- drive_download(
  as_id("1-wlFUkzEwWX1afRWLJlY_KEJs7BfsZim"), path = temp, overwrite = TRUE)
# Setting the date file folder
out <- unzip(temp, exdir = data_folder_Sample)
# Date files for normal processing example are deposited below
out
```
## QC Data Inspection
Look for significant contaminants using 3D graph plotting rentention time
and M/Z. Chunk should provide the RT range and MZ range in console.

``` {r data inspection}
# Inspect the MS data via a 3D image. "res" are used to specify the resolution for the MS data.
PerformDataInspect(data_folder_QC,res = 50)

```

## ROI Extractions
Software includes three methods for ROI extraction: 
1. Standard simulation method ("ssm_trim") which is the default
This method trims from m/z dinmension first and then RT dimension
2. RT dimension method ("rt_specific")
This method extracts + or - val from parts of the spectra
3. M/Z dimension method ("mz_specific")
This method extracts + or - val from parts of the spectra

The output file is a trimmed file saved as a .mzML

``` {r ROI}
# QC samples are trimmed with "ssm_trim" strategy. 
# The percentage of RT dimension retained is set as 20%.
raw_data <- PerformROIExtraction(data_folder_QC, rt.idx = 0.2, rmConts = FALSE)
```

## Initialize Parameters
Select the platform being used from the following list: UPLC-Q/E, UPLC-Q/TOF, 
UPLC-T/TOF, UPLC-Ion_trap, UPLC-Orbitrap, UPLC-G2S, HPLC-Q/TOF, HPLC-Ion_Trap, 
HPLC-Orbitrap and HPLC-S/Q. Specific parameters can be set manually as well.

``` {r param setting}
# Initial platform specific parameters
param_initial <- SetPeakParam(platform = "UPLC-Q/E") 
```


## Parameter Optimization
Perform parameter optimization using previously set parameters, input data, and
number of cores available. Chunk will calculate noise in ppm, spit the data, 
and iterate model parsing to increase Gaussian peak ratio

DON'T RERUN UNLESS YOU HAVE 330 MIN TO SPARE

``` {r parm optimize}
# Select relative core according to your work platform
param_optimized <- PerformParamsOptimization(raw_data, param = param_initial, ncore = 8)
```



## Peak Profiling
Reads the raw MS data files and visualizes as Total Ion Chromatogram (TIC) and
Base Peak Chromatogram (BPC). If a specific peak of interest wants to be
visualized, an Extracted Ion Chromatogram can be generated. 

``` {r peak profile}
# Import raw MS data. The "SetPlotParam" parameters can be used to determine plot or not.
rawData <- ImportRawMSData(NULL, data_folder_Sample, plotSettings = SetPlotParam(Plot=FALSE))
```


The optimized parameters previously found are then used to perform peak
detection, alignment, and grouping. The output of this function are two 
diagnostic plots (RT adjustment map and PCA plot showing sample clustering
prior to data cleaning and stat analysis) as well as statistics on the total
intensity of peaks from different samples.

``` {r run profile}
# Peak Profiling with optimized parameters
mSet <- PerformPeakProfiling(rawData,param_optimized$best_parameters,
                             plotSettings = SetPlotParam(Plot = T))
```


## Peak Annotation
Annotate isotopes and adduct peaks using a method from CAMERA. Output is a .csv
of annotated peaks and a formatted list of peaks for future use filtered based
on user specs.

``` {r annotation params}
# Setting the Annotation Parameters.
annParams <- SetAnnotationParam(polarity = "negative", mz_abs_add = 0.005)
```

``` {r peak annotation}
# Perform peak annotation.
annotPeaks <- PerformPeakAnnotation(mSet, annParams)
```

``` {r peak formatting}
## Format and filter the peak list for MetaboAnalystR
maPeaks <- FormatPeakList(annotPeaks, annParams, filtIso =F, filtAdducts = FALSE, missPercent = 1)
```

## ~~~~
## Data Processing and Statistical Analysis
## ~~~~

## Initialize data set

Need to initialize some objects will be used later down the line
``` {r init objects}
# First step is to create the mSet Object, specifying that the data to be uploaded
# is a peak table ("pktable") and that statistical analysis will be performed ("stat").
mSet<-InitDataObjects("pktable", "stat", FALSE)
```
Import saved data
```{r import filtered data}
# Second step is to read in the filtered peak list, please set the path right first
#mSet<-Read.TextData(mSet, "metaboanalyst_input.csv", "colu", "disc")
mSet<- maPeaks
```


Perform sanity checks to make sure that the data looks good before proceeding
```{r data check}
# The third step is to perform data processing using MetaboAnalystR (filtering/normalization)
# Perform data processing - Data checking
mSet<-SanityCheckData(mSet)
```

## Data Stats
???Why are we replacing the minimum values???

```{r min val}
# Perform data processing - Minimum Value Replacing
mSet<-ReplaceMin(mSet);
```






